---
title: "Expt1_food"
format: html
editor_options: 
  chunk_output_type: console
---

# Project information

Title: Optimization of grow‐out culture of Caribbean king crab: experiment 1 (food)
Funder: Gulf States Marine Fisheries Commission 
Award number: NA20NMF4540293 (subcontract ACQ-210-039-2023-USF)

PI(s): Christopher Stallings and Fredrick Stengard

Code developer(s): Christopher Stallings and Fredrick Stengard 
Origination Date: 2025-05-16

# Packages and parameters
```{r}
# Load required packages
library(tidyverse)  # for data manipulation and visualization
library(patchwork) # for arranging plots

# Import data
source("00_CommonFiles/DataImport.R")
```

# Visualization
## Change in size
```{r}

# Function for carapace width plots
plot_cw <- function(df, df_title) {
  df <- df %>% drop_na(cw_diff, cw_percent, ttt, sex)

  p1 <- ggplot(df, aes(x = ttt, y = cw_diff)) +
    geom_boxplot() +
    labs(title = df_title, x = "ttt", y = "Change in carapace width (mm)")

  p2 <- ggplot(df, aes(x = sex, y = cw_diff)) +
    geom_boxplot() +
    labs(title = df_title, x = "sex", y = "Change in carapace width (mm)")

  p3 <- ggplot(df, aes(x = ttt, y = cw_percent)) +
    geom_boxplot() +
    labs(title = df_title, x = "ttt", y = "Percent change in carapace width")

  p4 <- ggplot(df, aes(x = sex, y = cw_percent)) +
    geom_boxplot() +
    labs(title = df_title, x = "sex", y = "Percent change in carapace width")

  (p1 | p2) / (p3 | p4)
}

# Function for mass plots
plot_mass <- function(df, df_title) {
  df <- df %>% drop_na(mass_diff, mass_percent, ttt, sex)

  p1 <- ggplot(df, aes(x = ttt, y = mass_diff)) +
    geom_boxplot() +
    labs(title = df_title, x = "ttt", y = "Change in mass (g)")

  p2 <- ggplot(df, aes(x = sex, y = mass_diff)) +
    geom_boxplot() +
    labs(title = df_title, x = "sex", y = "Change in mass (g)")

  p3 <- ggplot(df, aes(x = ttt, y = mass_percent)) +
    geom_boxplot() +
    labs(title = df_title, x = "ttt", y = "Percent change in mass")

  p4 <- ggplot(df, aes(x = sex, y = mass_percent)) +
    geom_boxplot() +
    labs(title = df_title, x = "sex", y = "Percent change in mass")

  (p1 | p2) / (p3 | p4)
}

# Carapace width plots
plot_cw(dat_expt1, "Experiment 1: food")
plot_cw(dat_expt2, "Experiment 2: photoperiod")
plot_cw(dat_expt3, "Experiment 3: space")

# Mass plots
plot_mass(dat_expt1, "Experiment 1: food")
plot_mass(dat_expt2, "Experiment 2: photoperiod")
plot_mass(dat_expt3, "Experiment 3: space")

```

## Proportion molt
```{r}

# Function to plot molt proportions by treatment
plot_molt_by_ttt <- function(df, df_title) {
  df <- df %>% drop_na(molt, ttt)

  ggplot(df, aes(x = ttt, y = molt)) +
    stat_summary(fun = mean, geom = "bar", fill = "darkgreen") +
    stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.2) +
    labs(
      title = df_title,
      x = "Treatment",
      y = "Proportion molted"
    ) +
    ylim(0, 1)
}

# Generate and display plots for each dataset
plot_molt_by_ttt(dat_expt1, "Experiment 1: food")
plot_molt_by_ttt(dat_expt2, "Experiment 2: photoperiod")
plot_molt_by_ttt(dat_expt3, "Experiment 3: space")


```

# Analyses
## Check distributions and assumptions of t-tools
```{r}
library(tidyverse)
library(car)

# Helper function to run Shapiro-Wilk and Levene's test
check_assumptions <- function(df, var) {
  cat("\n---", var, "---\n")
  print(shapiro.test(df[[var]]))
  print(leveneTest(as.formula(paste(var, "~ ttt")), data = df))
}

# Helper function to run t-test or ANOVA
run_test <- function(df, var) {
  if (n_distinct(df$ttt) == 2) {
    print(t.test(as.formula(paste(var, "~ ttt")), data = df))
  } else {
    print(summary(aov(as.formula(paste(var, "~ ttt")), data = df)))
  }
}

# Variables to test
vars <- c("cw_diff", "cw_percent", "mass_diff", "mass_percent")

# dat_expt1: t-tests
df1 <- dat_expt1 %>% drop_na(ttt, cw_diff, cw_percent, mass_diff, mass_percent)
cat("=== Experiment 1: food ===\n")
for (v in vars) {
  check_assumptions(df1, v)
  run_test(df1, v)
}

# dat_expt2: ANOVAs
df2 <- dat_expt2 %>% drop_na(ttt, cw_diff, cw_percent, mass_diff, mass_percent)
cat("\n=== Experiment 2: photoperiod ===\n")
for (v in vars) {
  check_assumptions(df2, v)
  run_test(df2, v)
}

# dat_expt3: ANOVAs
df3 <- dat_expt3 %>% drop_na(ttt, cw_diff, cw_percent, mass_diff, mass_percent)
cat("\n=== Experiment 3: space ===\n")
for (v in vars) {
  check_assumptions(df3, v)
  run_test(df3, v)
}

```

## Assumptions take 2
```{r}

```
## Change in size
```{r}

```

## Proportion molt
```{r}

```
