---
title: "report_edits"
format: html
editor: visual
---

# Project information

Title: Optimization of grow‐out culture of Caribbean king crab Funder: Gulf States Marine Fisheries Commission Award number: NA20NMF4540293 (subcontract ACQ-210-039-2023-USF)

PI(s): Christopher Stallings and Fredrick Stengard

Code developer(s): Christopher Stallings Origination Date: 2025-07-31

This code is to fix issues we were having with several plots made in the main qmd file.

# Packages and parameters

```{r}
# Load required packages
library(tidyverse)  # for data manipulation and visualization
library(patchwork) # for arranging plots
library(dplyr) # for data manipulation
library(car) # for Levene's Test
library(ResourceSelection)  # for Hosmer-Lemeshow test
library(pscl)               # for pseudo R²
library(pROC)               # for ROC curve and AUC
library(ggplot2)

setwd("C:/Users/stallings/Box/1-Professional/Students/1a - Active Fish Lab/2. Stengard/1. Stengard_SWC/Funding/SCMFC 2022/CKC_Experiments")
# Import data
source("00_CommonFiles/DataImport.R")
```

#### Visualization: Change in size and time to molt

```{r}
# Function for carapace width plots
plot_cw <- function(df, df_title) {
  df <- df %>% drop_na(cw_diff, cw_percent, ttt, sex)

  p1 <- ggplot(df, aes(x = ttt, y = cw_diff)) +
    geom_boxplot() +
    labs(title = df_title, x = "ttt", y = "Change in carapace width (mm)")

  p2 <- ggplot(df, aes(x = sex, y = cw_diff)) +
    geom_boxplot() +
    labs(x = "sex", y = "Change in carapace width (mm)")

  p3 <- ggplot(df, aes(x = ttt, y = cw_percent)) +
    geom_boxplot() +
    labs(x = "ttt", y = "Percent change in carapace width")

  p4 <- ggplot(df, aes(x = sex, y = cw_percent)) +
    geom_boxplot() +
    labs(x = "sex", y = "Percent change in carapace width")

  (p1 | p2) / (p3 | p4)
}

# Function for mass plots
plot_mass <- function(df, df_title) {
  df <- df %>% drop_na(mass_diff, mass_percent, ttt, sex)

  p1 <- ggplot(df, aes(x = ttt, y = mass_diff)) +
    geom_boxplot() +
    labs(title = df_title, x = "ttt", y = "Change in mass (g)")

  p2 <- ggplot(df, aes(x = sex, y = mass_diff)) +
    geom_boxplot() +
    labs(x = "sex", y = "Change in mass (g)")

  p3 <- ggplot(df, aes(x = ttt, y = mass_percent)) +
    geom_boxplot() +
    labs(x = "ttt", y = "Percent change in mass")

  p4 <- ggplot(df, aes(x = sex, y = mass_percent)) +
    geom_boxplot() +
    labs(x = "sex", y = "Percent change in mass")

  (p1 | p2) / (p3 | p4)
}

# Function for time to molt plots
plot_time_to_molt <- function(df, df_title) {
  df <- df %>%
    drop_na(time_to_molt, ttt, sex)

  p1 <- ggplot(df, aes(x = ttt, y = time_to_molt)) +
    geom_boxplot() +
    labs(title = df_title, x = "ttt", y = "Time to molt (days)")

  p2 <- ggplot(df, aes(x = sex, y = time_to_molt)) +
    geom_boxplot() +
    labs(x = "sex", y = "Time to molt (days)")

  p1 | p2
}

# Generate plots
plot_cw(trial1_size, "Trial 1: food")
plot_cw(trial2_size, "Trial 2: photoperiod")
plot_cw(trial3_size, "Trial 3: space")

plot_mass(trial1_size, "Trial 1: food")
plot_mass(trial2_size, "Trial 2: photoperiod")
plot_mass(trial3_size, "Trial 3: space")

plot_time_to_molt(trial1_ttm, "Trial 1: food")
plot_time_to_molt(trial2_ttm, "Trial 2: photoperiod")
plot_time_to_molt(trial3_ttm, "Trial 3: space")

# Save plots
ggsave("03_Figures/cw_diff_trial1.png", plot_cw(trial1_size, "Trial 1: food"), width = 10, height = 8, dpi = 300)
ggsave("03_Figures/cw_diff_trial2.png", plot_cw(trial2_size, "Trial 2: photoperiod"), width = 10, height = 8, dpi = 300)
ggsave("03_Figures/cw_diff_trial3.png", plot_cw(trial3_size, "Trial 3: space"), width = 10, height = 8, dpi = 300)

ggsave("03_Figures/mass_diff_trial1.png", plot_mass(trial1_size, "Trial 1: food"), width = 10, height = 8, dpi = 300)
ggsave("03_Figures/mass_diff_trial2.png", plot_mass(trial2_size, "Trial 2: photoperiod"), width = 10, height = 8, dpi = 300)
ggsave("03_Figures/mass_diff_trial3.png", plot_mass(trial3_size, "Trial 3: space"), width = 10, height = 8, dpi = 300)

ggsave("03_Figures/time_to_molt_trial1.png", plot_time_to_molt(trial1_ttm, "Trial 1: food"), width = 8, height = 6, dpi = 300)
ggsave("03_Figures/time_to_molt_trial2.png", plot_time_to_molt(trial2_ttm, "Trial 2: photoperiod"), width = 8, height = 6, dpi = 300)
ggsave("03_Figures/time_to_molt_trial3.png", plot_time_to_molt(trial3_ttm, "Trial 3: space"), width = 8, height = 6, dpi = 300)
```

#Analyses \## Check distributions and assumptions of parametric t-tools

```{r}
library(car)  # for Levene's Test

# Function to check assumptions and save results
check_assumptions <- function(data, group_var, test_vars, dataset_name) {
  results <- list()
  
  for (var in test_vars) {
    var_results <- list()
    
    # Normality check using by()
    normality <- by(data[[var]], data[[group_var]], function(x) {
      x <- na.omit(x)
      group_name <- attr(x, "subset")
      if (length(x) >= 3) {
        shapiro <- shapiro.test(x)
        return(list(group = group_name, p_value = shapiro$p.value))
      } else {
        return(list(group = group_name, p_value = NA, note = "Insufficient data"))
      }
    })
    var_results$normality <- normality
    
    # Levene's Test
    formula <- as.formula(paste(var, "~", group_var))
    levene <- leveneTest(formula, data = data)
    var_results$levene <- capture.output(print(levene))
    
    results[[var]] <- var_results
  }
  
  # Prepare output text
  output <- c(paste0("=== Assumption Checks for ", dataset_name, " ===\n"))
  for (var in names(results)) {
    output <- c(output, paste0("\n--- ", var, " ---\n"))
    
    output <- c(output, "Shapiro-Wilk Normality Test:")
    for (res in results[[var]]$normality) {
      if (!is.null(res$note)) {
        output <- c(output, paste("Group:", res$group, "-", res$note))
      } else {
        output <- c(output, paste("Group:", res$group, "- p-value:", round(res$p_value, 5)))
      }
    }
    
    output <- c(output, "\nLevene's Test for Homogeneity of Variance:")
    output <- c(output, results[[var]]$levene)
  }
  
  return(output)
}

# Variables to test
size_vars <- c("cw_diff", "cw_percent", "mass_diff", "mass_percent")
ttm_var <- c("time_to_molt")

# Run checks
results1_size <- check_assumptions(trial1_size, "ttt", size_vars, "trial1_size")
results2_size <- check_assumptions(trial2_size, "ttt", size_vars, "trial2_size")
results3_size <- check_assumptions(trial3_size, "ttt", size_vars, "trial3_size")

results1_ttm <- check_assumptions(trial1_ttm, "ttt", ttm_var, "trial1_ttm")
results2_ttm <- check_assumptions(trial2_ttm, "ttt", ttm_var, "trial2_ttm")
results3_ttm <- check_assumptions(trial3_ttm, "ttt", ttm_var, "trial3_ttm")

# Combine all results
all_results <- c(
  results1_size, "\n\n", results2_size, "\n\n", results3_size, "\n\n",
  results1_ttm, "\n\n", results2_ttm, "\n\n", results3_ttm
)

# Create output directory if it doesn't exist
output_dir <- "02_Results"
if (!dir.exists(output_dir)) dir.create(output_dir)

# Write to file
writeLines(all_results, file.path(output_dir, "growth_diagnostics.txt"))
```

## Change in size & time to molt

```{r}
# Function to run parametric or non-parametric tests based on assumptions
run_mixed_tests <- function(data, group_var, test_vars, dataset_name, assumptions) {
  results <- list()
  
  for (var in test_vars) {
    df <- data[, c(group_var, var)]
    df <- na.omit(df)
    n_levels <- length(unique(df[[group_var]]))
    
    # Determine test type from assumptions
    use_parametric <- assumptions[[dataset_name]][[var]]$parametric
    
    if (n_levels < 2) {
      test_output <- "Not enough groups to run test"
    } else if (use_parametric) {
      if (n_levels == 2) {
        test <- t.test(as.formula(paste(var, "~", group_var)), data = df)
        test_output <- capture.output(print(test))
      } else {
        test <- aov(as.formula(paste(var, "~", group_var)), data = df)
        test_output <- capture.output(summary(test))
      }
      header <- paste0("--- ", var, " (Parametric) ---")
    } else {
      if (n_levels == 2) {
        test <- wilcox.test(as.formula(paste(var, "~", group_var)), data = df)
      } else {
        test <- kruskal.test(as.formula(paste(var, "~", group_var)), data = df)
      }
      test_output <- capture.output(print(test))
      header <- paste0("--- ", var, " (Non-Parametric) ---")
    }
    
    results[[var]] <- c(header, test_output)
  }
  
  output <- c(paste0("=== Test Results for ", dataset_name, " ===\n"))
  for (var in names(results)) {
    output <- c(output, results[[var]])
  }
  return(output)
}

# Define which variables meet parametric assumptions
assumptions <- list(
  trial1_size = list(
    cw_diff = list(parametric = FALSE),
    cw_percent = list(parametric = TRUE),
    mass_diff = list(parametric = FALSE),
    mass_percent = list(parametric = FALSE)
  ),
  trial2_size = list(
    cw_diff = list(parametric = TRUE),
    cw_percent = list(parametric = TRUE),
    mass_diff = list(parametric = FALSE),
    mass_percent = list(parametric = FALSE)
  ),
  trial3_size = list(
    cw_diff = list(parametric = FALSE),
    cw_percent = list(parametric = FALSE),
    mass_diff = list(parametric = TRUE),
    mass_percent = list(parametric = TRUE)
  ),
  trial1_ttm = list(time_to_molt = list(parametric = TRUE)),
  trial2_ttm = list(time_to_molt = list(parametric = FALSE)),
  trial3_ttm = list(time_to_molt = list(parametric = TRUE))
)

# Variables to test
size_vars <- c("cw_diff", "cw_percent", "mass_diff", "mass_percent")
ttm_var <- c("time_to_molt")

# Run tests and collect results
results1_size <- run_mixed_tests(trial1_size, "ttt", size_vars, "trial1_size", assumptions)
results2_size <- run_mixed_tests(trial2_size, "ttt", size_vars, "trial2_size", assumptions)
results3_size <- run_mixed_tests(trial3_size, "ttt", size_vars, "trial3_size", assumptions)

results1_ttm <- run_mixed_tests(trial1_ttm, "ttt", ttm_var, "trial1_ttm", assumptions)
results2_ttm <- run_mixed_tests(trial2_ttm, "ttt", ttm_var, "trial2_ttm", assumptions)
results3_ttm <- run_mixed_tests(trial3_ttm, "ttt", ttm_var, "trial3_ttm", assumptions)

# Combine all results
all_results <- c(
  results1_size, "\n\n", results2_size, "\n\n", results3_size, "\n\n",
  results1_ttm, "\n\n", results2_ttm, "\n\n", results3_ttm
)

# Create output directory if it doesn't exist
output_dir <- "02_Results"
if (!dir.exists(output_dir)) dir.create(output_dir)

# Write to file
writeLines(all_results, file.path(output_dir, "growth_results.txt"))
```
